---
title: "Cryoablation versus Thoracic Epidural for Pain Control Following Bilateral Lung Transplantation"
author: "Joy Chen"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output: bookdown::html_document2
---

```{r}
here::i_am('rmd/hw5.rmd')
```

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```
## Introduction

## Methods

```{r, load-data}
lungtx<-read.csv("cryo_dataset.csv")
```
We retrospectively identified a cohort of `r nrow(lungtx)` patients who underwent bilateral lung transplantation at our center between May 2016-August 2018. Patients requiring re-transplantation or extracorporeal membrane oxygenation prior to transplantation were excluded. Thoracic epidural catheter was placed on the first postoperative day in all patients 
(N=`r (sum(lungtx$group==1))`) prior to July 2017, and Cryo was utilized on all subsequent patients (N=`r (sum(lungtx$group==2))`). In regard to the Cryo technique, bilateral interspaces 3-7 were ablated with the cyroICE probe (AtriCure Inc; Mason, OH) during surgery under direct visualization. Postoperative multimodality analgesia was performed per standard of care. The primary outcome was total oral morphine equivalent (OME) usage through postoperative day (POD) 7 and secondary outcomes were pain scores through POD7, analgesic-specific complications, ambulation, and length of stay.

## Results

In lung transplantation, Cryoabalation was safe and had similar analgesic efficacy to thoracic epidural, as seen by similar pain scores on postoperative days 1 through 7 (Figure \@ref(fig:pain-plot).  

```{r, cleaning}
library(dplyr)
library(tidyr)
library(ggplot2)

lungtx <- lungtx %>% rowwise() %>% 
  mutate(podavg1=mean(c(pod16:pod124), na.rm=T)) %>%
  mutate(podavg2=mean(c(pod26:pod224), na.rm=T)) %>%
  mutate(podavg3=mean(c(pod36:pod324), na.rm=T)) %>%
  mutate(podavg4=mean(c(pod46:pod424), na.rm=T)) %>%
  mutate(podavg5=mean(c(pod56:pod524), na.rm=T)) %>%
  mutate(podavg6=mean(c(pod66:pod624), na.rm=T)) %>%
  mutate(podavg7=mean(c(pod76:pod724), na.rm=T)) 

pain <- subset(lungtx, select=c(id, group, podavg1:podavg7))
pain<-pain %>%
  pivot_longer(
    cols = starts_with("podavg"),
    names_to = "POD",
    values_to = "Score",
  )

pain$POD[pain$POD == 'podavg1']<-1
pain$POD[pain$POD == 'podavg2']<-2
pain$POD[pain$POD == 'podavg3']<-3
pain$POD[pain$POD == 'podavg4']<-4
pain$POD[pain$POD == 'podavg5']<-5
pain$POD[pain$POD == 'podavg6']<-6
pain$POD[pain$POD == 'podavg7']<-7
pain$POD<-as.numeric(pain$POD)
pain$group<-as.factor(pain$group)
```

```{r, pain-plot}
ggplot(pain, aes(x=POD, y = Score, color = group)) +
  geom_point() +
  geom_smooth() +
  labs(x="Post Operative Day", y = "Pain Scale 0-10")
```

## Discussion
